---
title: "post02-meiying-huang"
author: "Meiying Huang"
date: "11/30/2017"
output: html_document
---
#Introduction: 
##On December 1, 1969, the United States held its first draft lottery ever since 1942, which gave young men (any year between 1944 and 1950 were assigned lottery) a random number corresponding to their birthdays. As part of the war in Vietnam, the United States government implemented a lottery to draft men of appropriate ages into the US Armed Forces. Under the old system of drafting first the oldest available men age 19 through 25, in 1970, a drawing took place in which capsules corresponding to each day of the year (numbered sequentially as 1-366) were placed into a drum and selected one-by-one. The order of selection determined the order in which men would be drafted by matching each eligible male to the capsule labeled with his birthday; hence, men whose birthdays were selected earlier would be drafted into the military earlier. 
##This post is focusing on analysing simulating randomization test for correlation as the following question:

#)How the draft worked? 
## The method of the draft lottery: the days of the year(including Feb 29) were written on slips of paper. These pieces of paper were then placed in separate plastic capsules that were mixed in a shoebox and then dumped into a deep glass jar. Capsules were drawn from the jar one at a time.

#If you had been born in 1950, you would be at least eighteen years old. Would your draft number have been called on December 1, 1969 among the first number and the second number being drew? You can test here to find out.
```{r}
# sicne lottery number 1 is sep 19 : 177
#m1969<- lottery$Month_Number=="9"
#d1969<- lottery$Day_of_year=="263"
m1=9
d1=19
m2=4
d2=24
test_my_bday<- function(m,d){
  if((m == m1)&(d== d1)){
    print("Oops, you was called in 1970 in the first two draws")
  }else if((m==m2) & (d==d2)){
    print("Oops, you was called in 1970 in the first two draws")
   } else{
    print("Lucky you! You was NOT called in 1970 in the first two draws")
    }
}

#let me test my birthday
test_my_bday(4,19)  
#go ahead to test your birthday here
# text_my_bday(m,d)

```
## According to Dan Eberhart's "A pact with living", during the first selective lottery, 85,000 American young man between the ages of 19 to 25. Are there any relationship between different variables?

#Data: According to the RESULTS FROM LOTTERY DRAWING - Vietnam Era 1970. You can find it on the file 1970 lottery. It is named "data_1970lottery.txt" and is vailable in the "post01." The columns of the file data_1970lottery have month, month number, day of teh eyar, and draft number. Here will consider two variables recorded on each date, nad sequential date in the year(e.g, Jan 31 transcribed into sequential date as 31, Feb 1 as 32, and December 31 as 366, and so on).

# Load the data files    ( Since this post is being asked to write it in a way which is also easy for the other audience with (almost) no experience in this area. And it is easier for people to run the code without worrying about the wrong paths, directory issue. I stored every file in post02 instead of putting more different files inside).
```{r}
lottery = read.table('data_1970lottery.txt', sep = ',', header = T)
```

##First, let's see if there is any relationship between the draft number and the day number by plotting the draft numver versus day number. In other word, in a perfectly fair, random lottery, what should be the value of the correlation coefficient between draft number and sequential date of birthday?


# - Scatterplot
```{r}
attach(lottery)
# Make a scatterplot of draft number against the day number to see if anything appear amiss?
#consider the day number as the independent varialbe x, and draft number is the dependent variable y.
plot(lottery$Day_of_year, lottery$Draft_No, main = "day number vs draft number", 
     xlab = "day number", ylab = "draft number", pch= 20)
# include a regression line
abline(lm(lottery$Draft_No ~ lottery$Day_of_year), col = "red")
#LOWESS smoother which uses locally-weighted polynomial regression give the coordinates of the smooth
lines(lowess(lottery$Day_of_year, lottery$Draft_No), col = "blue")

##From the above scatterplot, we observe that the data points are scattered randomly. The plot shows there seems has no relationship between the draft numbers and the day numbers. Therefore we conclude that the day numbers and the draft numbers are independent. The scatter plot does not reval much of an association between draft number and sequential dates.
```

```{r}
#Find the correlations: Compute both the Pearson and Spearman correlations between day number and draft number. Compare their values. Does this make sense? Return: Pearson and Spear- man correlations.

#find Pearson correlations between day number and draft number.
##draft_pearson_cor
draft_pearson_cor<- cor.test(lottery$Day_of_year, 
                             lottery$Draft_No,
                              method = "pearson")
draft_pearson_cor

# find Spearman correlations between day number and draft number.
##draft_spearman_cor
draft_spearman_cor<-cor.test(lottery$Day_of_year,
                             lottery$Draft_No,
                              method =  "spearman")
draft_spearman_cor

obs_corr<- cor(lottery$Day_of_year, lottery$Draft_No)
##result: the Pearson and Spearman correlations between day number and draft number are the same -0.2256861. Which means x(day number) and y(draft number) has a week negative linear relationship, means when one variable tends to increase as the other decreases. The correlation shows us that the highest value of draft number is tend to associated with the lowest value of day number but the relationship is weak. 
```


```{r}
##To determine whether the observed correlation is statistically significant, we need to perform a permutation test with a big enough sample size ( I did it with 10,000 permutations). 
#H0(null hypothesis): p = -0.2260414, H1:(alternative hypothesis) p > -0.2260414
draft_num<- lottery$Draft_No
day_num<- lottery$Day_of_year
corr.random =  NULL
for(i in 1:10000){
  draft_random = sample(draft_num, length(draft_num), replace = TRUE )
  corr.random[i] = cor(day_num, draft_random)
}
p_value= sum(corr.random > obs_corr)/10000
p_value


#Provide a histogram of the correlation values arising from your permutations and annotate the true correlation. Compute the empirical p-value. Is the correlation significant? Return: empirical p-value.
hist(corr.random, main = "a histogram of the correlation values")
abline(v= obs_corr, col = "red", lwd = 2) 
#result: assuming significance level=5%, since here p_value=1 which is greater than 5%, so fail to reject the H0(null hypothesis). Thus, the observed correlation is statistically significant.

```

```{r}
#plot a boxplot of draft number against months to see (if any) what trend here has?
month_num<- lottery$Month_Number
boxplot(draft_num ~ month_num,
        main = "draft number against months",
        xlab =  "months", ylab = "draft number")
##December has the losest median draft number and has less the lowest minimum and lowest maximum draft number. It's more likely to have an average around 200 draft number. 
##The average draft number per month increases from January to March, and from August to October. Decreases from March to August, and from October to December. There's no clear increasing or decreasing pattern through out the months in a year.
```

```{r}
#Why is the Kruskal-Wallis test more appropriate than ANOVA for this data? Perform the Kruskal-Wallis test to see if draft numbers are different across months and describe what you would conclude based on the result. Return: KW test p-value.
kruskal.test(lottery$Month ~ draft_num)
kw_p_val = 0.4902
```
```{r}
#How many pairwise comparisons are possible? For each comparison, perform a Mann- Whitney U test. Return: the set of (uncorrected) p-values.
#since we have n=366, comparison=n(n-1)/2=66,795 with the pairwise comparison method.
# a helpful function is combn
mw_p_vec1<- wilcox.test(draft_num, month_num, correct = FALSE)
mw_p_vec1
## p-value1 < 2.2e-16, which is less than 0.05. So reject the null hypothesis, means that these two groups(draft number and month number do not have statistical equality mean)
sum.rank.draft = sum(rank(c(draft_num, month_num)) [1:366])
w1=sum.rank.draft- length(draft_num)*(length(draft_num)+1)/2
w1
sum.rank.month = sum(rank(c(draft_num, month_num)) [367: 732])
w2 = sum.rank.month -  length(month_num)*(length(month_num)+1)/2
w2
w3= sum(rank(c(draft_num,month_num))[1:366])
w3
w4 = sum(rank(c(draft_num,month_num))[367:732])
w4  
##in the group(draft and month number) since the compute interval(w3=198916, w4=69362) is contained within the tabulated interval(w1=131755, w2= 2201), we conclude by accepting the hypothesis H0 of equality of means.

mw_p_vec2<- wilcox.test(draft_num, day_num, correct = FALSE)
mw_p_vec2
##p-value = 1, greater than 0.05. then we can accepct the null hypothesis H0 of statistical equality of the means of two group(draft number and day number.)
sum.rank.draft2 = sum(rank(c(draft_num, day_num)) [1:366])
w11=sum.rank.draft2- length(draft_num)*(length(draft_num)+1)/2
w11
sum.rank.day = sum(rank(c(draft_num, day_num)) [367: 732])
w22 = sum.rank.day -  length(day_num)*(length(day_num)+1)/2
w2
w33= sum(rank(c(draft_num,day_num))[1:366])
w33
w44 = sum(rank(c(draft_num,day_num))[367:732])
w44 
```


```{r}
# Which comparisons are significant using the Bonferroni (FWER) correction with α = 0.05? What about the Benjamini-Hochberg (FDR) correction at α = 0.05? Return: the set of corrected p-values for both methods.
kw_p_val = 0.4902
p1 = kw_p_val
fwer_p_vec1<- p.adjust(p1, method = "bonferroni", n= length(p1))
fwer_p_vec1

fdr_p_vec1<- p.adjust(p1, method = "fdr", n= length(p1))
fdr_p_vec1

p2 = 1
fwer_p_vec2<- p.adjust(p2, method = "bonferroni", n= length(p2))
fwer_p_vec2

fdr_p_vec2<- p.adjust(p2, method = "fdr", n= length(p2))
fdr_p_vec2

```

```{r}
#Perform Fisher’s exact test with the following categories in the table: January-June vs July-December and draft numbers of 1-182 vs 183-366. Provide the table and compare the p-value to that from the Kruskal-Wallis test. Return: p-value.
m1to6= c(lottery$Month_Number<7, lottery$Month_Number) 
m1to6
sum(m1to6)

m7to12= c(lottery$Month_Number>6, lottery$Month_Number) 
m7to12
sum(m7to12)

low_draft= c(lottery$Draft_No< 183, lottery$Draft_No)
low_draft
length(low_draft)

high_draft= c(lottery$Draft_No> 182, lottery$Draft_No)
high_draft
length(high_draft)

m1to6_lowdraft= c(m1to6[draft_num< 183])
a11= sum(m1to6_lowdraft)
a11
m1to6_highdraft= c(m1to6[draft_num> 182])
a12= sum(m1to6_highdraft)
a12

m7to12_lowdraft= c(m7to12[draft_num< 183])
a21= sum(m7to12_lowdraft)
a21
m7to12_highdraft= c(m7to12[draft_num> 182])
a22= sum(m7to12_highdraft)
a22

p1=a11/(a11+a12)
p1

p2= a21/(a21+a22)
p2

test1<- matrix(c(a11,a12,a21,a22), nrow = 2,
               dimnames= list(month= c("1to6", "7to12"),
                              draft= c("low", "high")))

fet_p_val<- fisher.test(test1, alternative = "two.sided")
fet_p_val

#result: Since p-value = 0.312, the p-value is not small enough to reject the null hypothesis,  conclusion will continue to hold even if the dependence is not properly accounted for. So what would you tell the congress when they call you to testify about the fairness of the lottery? What are the advantages and disadvantages of the Kruskal-Wallis test against Fisher’s exact test in this case?
#
```


```{r}
#for further research:
#The beauty of the 1970 draft lottery data is that students can confirm the nonrandomness in the lottery process by a wide variety of approaches. The dataset is ideal for computer laboratory experiments and for graphical exploration. A few Minitab examples will illustrate the ease with which one can display the lottery's bias. Also analyzed are the corresponding data for the 1971 lottery, which featured a much improved, if more complicated, randomization. (The 1970 draft lottery applied to eligible men aged 19 to 26 prior to January 1, 1970, and so included births taking place in some leap years. The 1971 draft lottery applied to men born in 1951, so only 365 days were involved.) 
```


```{r}
#Reference
#1,Larsen, R. J., and Stroup, D. F. (1976), Statistics in the Real World, New York: Macmillan, pp. 241-245. Nonparametric approaches: Kruskal-Wallis test on the monthly ranks and Wilcoxon/Mann-Whitney rank sum test on the ranks in the first and last halves of the year.

#2, Landscaper.net (2009). The Military Draft and 1969 Draft Lottery for the Vietnam War. Last modified 2009-03-24. Confirmed 2011-05-26. 

#3, "The vietnam lotteries" Last Updated June 18, 2009 https://web.archive.org/web/20120915225120/https://www.sss.gov/lotter1.htm

#4, RESULTS FROM LOTTERY DRAWING - Vietnam Era 1970 #https://web.archive.org/web/20120915232321/http://www.sss.gov/LOTTER8.HTM

#5, Eberhart, Dan. A pact with the living. 10-26-2016. Print

#6, Draft lottery (1969) https://en.wikipedia.org/wiki/Draft_lottery_(1969)#cite_note-SSS70-1

#7, “Would your draft number have been called?” USA Today, Gannett Satellite Information Network, 2017, www.usatoday.com/vietnam-war/draft-picker.

#8,Rosenbaum, D. E. (1970a), "Statisticians Charge Draft Lottery Was Not Random," New York Times, January 4, p. 66. "If the results occur less frequently" than 5% of the time, "then the statisticians conclude that some causative factor was involved." This makes an interesting contrast to the subsequent delicacy with which the New York Times characterizes the statistics involved in polls ("How the Poll Was Conducted").
```

